---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    self_contained: no
  html_notebook:
    self_contained: no
editor_options:
  chunk_output_type: console
---


```{r, eval=FALSE}
rm(list = ls()) # I don't care what you think
for(i in dev.list()) dev.off()
setwd("scripts/")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Old Source of the criblage data:  
<https://www.data.gouv.fr/fr/datasets/donnees-de-laboratoires-pour-le-depistage-indicateurs-sur-les-mutations/>

New Source: 
<https://www.data.gouv.fr/fr/datasets/r/3e4fa086-8642-4d30-8316-7d4ec4cf893b>
# Initializations

```{r loadClearMergeData}
# Whether to download the data
# (if cleanMergeData == TRUE only)
dlData <- FALSE

# Whether to recompute the combined datasets and delissage
cleanMergeData <- FALSE
if(dlData) cleanMergeData <- TRUE
```

# Load data 

```{r, results = 'hide'}
if(cleanMergeData){
  # Compute datasets
  source("load-clean-merge-data.R")
}

# Load datasets
dat.France <- read.csv("../data/datCribTest_France.csv")
dat.Regions <- read.csv("../data/datCribTest_Regions.csv")
dat.Deps <- read.csv("../data/datCribTest_Deps.csv")


# Turn dates back into dates
dat.France$date <- as.Date(dat.France$date)
dat.Regions$date <- as.Date(dat.Regions$date)
dat.Deps$date <- as.Date(dat.Deps$date)

```

## Other

```{r}
# Define inverse logit
myinvlogit <- function(x) exp(x) / (1 + exp(x))

# Logit
logit <- function(p){
  log(p / (1-p))
} 
```

# Plots

## Parameters

```{r}
library(MetBrewer)
pal <- met.brewer("Hokusai3", 2, "discrete")
colLiss <- pal[2]
colDeliss <- pal[1]

pal <- MetBrewer::met.brewer("Java", 5, "discrete")
colLiss2 <- pal[1]
colDeliss2 <- pal[4]
```

## Define plotting function (and plot France)

```{r}

yvalues <- c(0.9999, 0.999, 0.99, 0.9, 0.75, 0.5, 0.25, 0.1, 0.05, 0.025, 0.01, 0.001, 0.0001)

plotMut0 <- function(time, test0, test1, col, thetit = "", ymin = 10^-6, ymax = 1 - 10^-6, denom = 7, scale = "logit", add = FALSE, addNotes = TRUE, pch = 16, lang = "FR", percent = TRUE, axes = TRUE, title = TRUE){
  # -- here showing n0/(n0+n1) --
  # time: vector of values of times of sample collection
  # test0: vector of numbers of tests with value 0
  # test1: vector of numbers of tests with value 1
  # col: color used for the plot
  # thetit: title of the plot
  # ymin: min y value
  # ymax: max y value
  # denom: denominator value; 7 for weekly sums, 1 for daily data
  # scale: "logit" or "linear"
  # addNotes: whether to show warning notes on the plot
  # addDeliss: whether to add data délissées (if available)
  # timedl: time values of the délissées data
  # test0dl: C0 values délissées
  # test1dl: C1 values délissées
  # coldl: color for the délissées data
  # addLegend: whether to add legend
  # add: if add to already existing plot
  
  n <- (test1 + test0) / denom
  p <- test0 / (test1 + test0)
  
  # Computation of the confidence interval
  deltaItv <- 1.96 * sqrt(p * (1-p) / n)
  keepPts <- !is.na(deltaItv) 
  # Remove points for which the interval cannot be computed
  deltaItv <- deltaItv[keepPts]
  pp <- p[keepPts]
  
  dxx <- 0 # Number of days to be added (depends on whether data délissées are used)
  
  cexL <- 1 # Point size of lissées data

  # Compute new values on the new scale
  changeScale <- function(p, confInt = FALSE){
    out <- NA # Initialize output, have NA if scale not properly written
    if(scale == "logit"){
      # Make sure that p is between 0 and 1 (can be different with conf int)
      p[p < ymin] <- max(0, 0.5*ymin)
      p[p > ymax] <- 2*ymax
      
      out <- log(p /(1-p)) # Logit scale
    }
    if(scale == "linear"){
      out <- p # Linear scale, no change
    }
    out
  }
  par(xpd = FALSE)
  par(las = 1, mgp = c(2, 0.3, 0), tck = -0.01)
  
  xmin <- min(as.Date(time), na.rm = TRUE)
  xmax <- max(as.Date(time), na.rm = TRUE)

  factor <- ifelse(percent, 100, 1)

  if(!add){
    # Create new plot
    plot(as.Date(time), changeScale(p), 
       ylim = c(changeScale(ymin), changeScale(ymax)), 
       xlim = c(xmin, xmax),
       frame.plot = FALSE, 
       xlab = "", ylab = "p", 
       col = col, pch = 16, 
       axes = FALSE, 
       yaxs = "i", type = "n")

      # Add graduations
      if(scale == "linear"){
        colGrad <- gray(0.85)
        for(i in seq(0, 1, by = 0.1)){
          abline(h = i, col = colGrad, lwd = 1.2)
        }
        for(i in seq(0.05, 0.95, by = 0.1)){
          #      abline(h = i, col = colGrad, lwd = 0.8)
        }
        abline(h = 0.5, col = colGrad, lwd = 3)
      }
      
    }  
  
  
  xx <- base::as.Date(time)[keepPts]

  # Add data points  
  points(as.Date(time), changeScale(p), 
       type = "p", col = col, pch = pch, cex = cexL)
  
  # Add confidence interval as arrows
  arrows(x0 = xx, 
         x1 = xx, 
         y0 = changeScale(pp - deltaItv), 
         y1 = changeScale(pp + deltaItv), 
         code = 0, 
         lwd = 3, 
         col = adjustcolor(col, alpha.f = 0.25), 
         lend = 1)


  if(addNotes){
  # Add note about data and code provenance
  mtext("Data: https://www.data.gouv.fr/fr/datasets/donnees-de-laboratoires-pour-le-depistage-indicateurs-sur-les-mutations/
Code: https://github.com/flodebarre/nouveauCriblage/blob/main/scripts/criblage2.Rmd", side = 1, line = 2.5, cex = 0.5, col = gray(0.5), adj = 0, family = "mono")
   
    
    if(lang == "FR"){
      txt <- c("À partir de données déjà lissées sur 7j (comme fournies)", "À partir de données délissées (approximatives)")
    }else{
     txt <- c("7-d averaged data (as provided)", "'un-smoothened' data") 
    }
    legend("topleft", bty = "n", col = c(colLiss, colDeliss), pch = c(16, 1), legend = txt) 
  }

  
  if(axes & !add){
    # Axes: 
    # horizontal
    xxl <- seq(as.Date(min(time)), as.Date(max(time)) + dxx, by = "day")
    axis(1, at = xxl, 
         labels = format(seq(as.Date(min(time)), as.Date(max(time)) + dxx, by = "day"), "%d/%m"), 
         las = 2, 
         cex.axis = 0.6, 
         lwd = 0)
    
    # Vertical axes
    if(scale == "linear"){
      axis(2, lwd = 0)
      axis(4, lwd = 0)
    }
    if(scale == "logit"){
      axis(2, at = changeScale(yvalues), labels = factor * yvalues)
      axis(4, at = changeScale(yvalues), labels = factor * yvalues)
    }
  }
  
  if(title){
    title(main = thetit)
  }
#  print(p[length(p)])
}
```

#### L452R / C

```{r L452R_France}
# Plot France
layout(1)
par(mar = c(4, 4, 4, 4))
ix <- which(dat.France$date >= "2022-03-01")
ym <- 5*10^-4 # plot range, minimal y value
yM <- 0.5 # Plot range, max y value
lang <- "EN"
percent <- FALSE
# Plot data as shared on data.gouv.fr
tt <- ifelse(percent, ifelse(lang == "FR", "Pourcentage", "Percentage"), "Proportion")
tit <- ifelse(lang == "FR", "Proportion de tests L452R+ (et Q?), France", "Proportion of L452R+ (and Q?) tests, France") # Title
plotMut0(dat.France[ix, "date"], dat.France[ix, "nb_C1.sum7_mid"], dat.France[ix, "nb_C0.sum7_mid"], col = colLiss, thetit = tit, ymin = ym, ymax = yM, lang = lang, percent = percent)
# Add data "delisse"
plotMut0(dat.France[ix, "date"], dat.France[ix, "nb_C1.dl"], dat.France[ix, "nb_C0.dl"], col = colDeliss, add = TRUE, pch = 1, addNotes = FALSE)
```

#### K417N / D

```{r D_France}
# Plot France
layout(1)
par(mar = c(4, 4, 4, 4))
ix <- which(dat.France$date >= "2022-01-01")
ym <- 0.9 # plot range, minimal y value
yM <- 1 - 10^-3 # Plot range, max y value
# Plot data as shared on data.gouv.fr
plotMut0(dat.France[ix, "date"], dat.France[ix, "nb_D1.sum7_mid"], dat.France[ix, "nb_D0.sum7_mid"], col = colLiss, thetit = "Proportion of K417N+ tests, France", ymin = ym, ymax = yM, percent = percent)
# Add data "delisse"
plotMut0(dat.France[ix, "date"], dat.France[ix, "nb_D1.dl"], dat.France[ix, "nb_D0.dl"], col = colDeliss, add = TRUE, pch = 1, addNotes = FALSE)
```

### Stats

```{r fitFrance}
dat <- dat.France # Dataset to be used
n1 <- "nb_C1.dl" # Numbers of + (or -)
n2 <- "nb_C0.dl" # Numbers of - (or +)
dx <- 30 # Number of days forward for projection
nDaysEstim <- 21 # Number of days back for the estimation

# Day at which estimation starts
first_date <- max(dat$date, na.rm = TRUE) - 3 - nDaysEstim

# Subset of the data after that date
sub <- dat[dat$date >= first_date, ]
# Define time column (obsolete)
sub$time <- as.numeric(sub$date - sub$date[1])

# Fit logistic model
glm0 <- glm(as.matrix(sub[, c(n1, n2)]) ~  date, family = binomial(link="logit"), data = sub)

# Print summary
print(summary(glm0))
glm0$coefficients

# x values for the prediction (dx days after the last one)
# xpred <- seq(min(sub$time), max(sub$time) + dx)
xpred <- seq(min(sub$date), max(sub$date) + dx, by = "day")
# Predicted values from the model
pred <- predict(glm0, type = "response", newdata = data.frame(date = xpred), se.fit = TRUE)
plot(xpred, pred$fit, type = "l", lwd = 2, 
     xlab = "date", ylab = "Proportion", axes = FALSE)
#pred2 <- predict(glm0, type = "link", newdata = data.frame(time = xpred), se.fit = TRUE)
#points(xpred, myinvlogit(pred2$fit), pch = 2, col = 2)

  # Plot the contour of the prediction as polygon +/- CI
#  polygon(x = c(xpred, rev(xpred)), y = c(myinvlogit(pred2$fit - 1.96 * pred2$se.fit), rev(myinvlogit(pred2$fit + 1.96 * pred2$se.fit))), col = adjustcolor("pink", 0.4), border = NA)
  polygon(x = c(xpred, rev(xpred)), y = c(pred$fit - 1.96 * pred$se.fit, rev(pred$fit + 1.96 * pred$se.fit)), col = adjustcolor(colDeliss, 0.4), border = NA)
  points(x = sub$date, y = sub[, n1] / (sub[, n1] + sub[, n2]), col = colDeliss)
axis(1, at = xpred, labels = format(xpred, "%d/%m"), las = 3, cex.axis = 0.9)
axis(2, at = seq(0, 1, by = 0.05), las = 1)

# Logit version  
plot(xpred, logit(pred$fit), type = "l", lwd = 2, 
     xlab = "date", ylab = "Proportion (logit)", axes = FALSE)
axis(1, at = xpred, labels = format(xpred, "%d/%m"), las = 3, cex.axis = 0.9)
axis(2, at = logit(yvalues), labels = yvalues, las = 1)
  
polygon(x = c(xpred, rev(xpred)), y = logit(c(pred$fit - 1.96 * pred$se.fit, rev(pred$fit + 1.96 * pred$se.fit))), col = adjustcolor(colDeliss, 0.4), border = NA)

points(x = sub$date, y = logit(sub[, n1] / (sub[, n1] + sub[, n2])), col = colDeliss)

xpred[which(pred$fit > 0.5)[1]]

```

```{r}
first_date
subb <- dat.France[which(dat.France$date >= "2022-05-01"), ]
subb <- head(subb, -3)
tail(subb)
yplus <- subb$P7j * subb$nb_C1.sum7_mid / (subb$nb_C1.sum7_mid + subb$nb_C0.sum7_mid)
plot(as.Date(subb$date), yplus, ylim = c(0, 10000))

xx <- as.Date(subb$date)
yminus <- subb$nb_C0.sum7_mid / (subb$nb_C1.sum7_mid + subb$nb_C0.sum7_mid) * subb$P7j
plot(xx, yminus, ylim = c(100, max(yminus, na.rm = TRUE)), log = "y")
points(xx, yplus)

logP <- log(yplus)
logQ <- log(yminus)

glmP <- glm(logP ~ xx)
glmQ <- glm(logQ ~ xx)

summary(glmP)
(glmP$coefficients)["xx"]
log(2)/(glmP$coefficients)["xx"]

# Predicted values
dx <- 30
xpred <- seq(as.Date(min(xx)), as.Date(max(xx)) + dx, by = "day")
predLogP <- predict(glmP, newdata = data.frame(xx = xpred), se.fit = TRUE)
predLogQ <- predict(glmQ, newdata = data.frame(xx = xpred), se.fit = TRUE)

par(mar = c(4, 4, 2, 1), 
    mgp = c(3, 0.5, 0))
colQD <- "#8E508E" # In triad of colDeliss in Paletton
plot(xpred, exp(predLogP$fit), type = "l", ylim = c(0, 50000), 
     xlab = "", ylab = "number of cases", 
     frame.plot = FALSE,
     yaxs = "i", xaxs = "i",
     main = "Toy model, France", 
     col = colDeliss, lwd = 2)
lines(xpred, exp(predLogQ$fit), col = colQD, lwd = 2)
lines(xpred, exp(predLogP$fit) + exp(predLogQ$fit), col = "black")
legend("topright", col = c(colDeliss, colQD, "black"), lwd = c(2, 2, 1), legend = c("452R+ (and Q?)", "452R-", "total cases"), bty = "n")
```

### History

```{r}
minDate <- "2021-05-01"
tmp <- dat.France[dat.France$date >= minDate, ]
par(las = 1)
par(mar = c(4, 4, 2, 1))
thecol <- "#FF7F50"
plot(tmp$date, tmp$nb_C1.sum7_mid / (tmp$nb_C0.sum7_mid + tmp$nb_C1.sum7_mid), type = "l", axes = FALSE, 
     xlab = "", ylab = "Proportion L452R", lwd = 3, col = thecol)
mnth <- seq(as.Date(minDate), Sys.Date(), by = "month")
axis(1, at = mnth, labels = format(mnth, "%b\n%Y"), padj = 0.5, pos = 0)
axis(2, at = seq(0, 1, by = 0.1))


text(x = as.Date("2021-07-15"), y = 0.2, labels = "1ere\nvague\nDelta", adj = 0.5, col = "#AF601A")

text(x = as.Date("2022-01-15"), y = 0.8, labels = "vague\nOmicron", col = "#1F618D")

text(x = as.Date("2022-05-01"), y = 0.2 , labels = "Sous-lignées\n Omicron")

mtext(text = 
"Données : https://www.data.gouv.fr/fr/datasets/donnees-de-laboratoires-
           pour-le-depistage-a-compter-du-18-05-2022-si-dep/", family = "mono", col = gray(0.5), cex = 0.7, side = 1, line = 2.75, adj = 0)
```


## By region 

```{r}
# Define function to plot a specific region reg
plotReg <- function(reg, xmin = "2022-04-01", addNotes = FALSE, ymax = 0.5, ymin = 1*10^-3, withMut = TRUE, col = colLiss, scale = "logit", percent = FALSE, lang = "FR", ...){
  # withMut: whether to plot C1 / (C1 + C0) (TRUE), or C0 / (C1 + C0)
  
  # Subset of the data for this region
  dat <- dat.Regions[which(dat.Regions$reg_name == reg & dat.Regions$date >= xmin), ]
  
  #suffix <- ifelse(deliss, ".dl", "")
  id <- ifelse(withMut, 1, 0)
  
  # Plot data as shared on data.gouv.fr
  suffix <- ".sum7_mid"
  plotMut0(dat$date, dat[, paste0("nb_C", toString(id), suffix)], dat[, paste0("nb_C", toString(1 - id), suffix)], col = colLiss, thetit = "", 
           ymin = ymin, ymax = ymax, 
           addNotes = addNotes, scale = scale, percent = percent, lang = lang, ...)
  
  # Plot data delisse
  suffix <- ".dl"
  plotMut0(dat$date, dat[, paste0("nb_C", toString(id), suffix)], dat[, paste0("nb_C", toString(1 - id), suffix)], col = colDeliss, thetit = "", 
           ymin = ymin, ymax = ymax, 
           addNotes = addNotes, add = TRUE, pch = 1, scale = scale, lang = lang, ...)

  # Region name as title  
  title(main = reg)
}

# Check function with IDF region
layout(1)
plotReg("Île-de-France")
```

```{r L452R_Regions, fig.width=7, fig.height=10}

# Multi-figure plot
layout(matrix(1:15, ncol = 3, byrow = TRUE))

# 1 Plot legend
par(mar = c(0, 0, 0, 0))
plot(0, 0, xlab = "", ylab = "", type = "n", axes = FALSE, xlim = c(-0.2, 1), ylim = c(-1, 0.5))
legend(x = -0.15, y = 0, legend = c("Données publiques, partagées déjà lissées", "Données dé-lissées"), 
          col = c(colLiss, colDeliss), 
          pch = c(16, 1), 
          box.col = gray(0, 0), bty = "n", cex = 0.9, pt.cex = 1.75, 
          title = "Proportion de mutation L452R (C1),\néchelle logit")

# 2 HDF
#layout(1)
themar <- c(2, 3, 3, 3)
par(mar = themar)
plotReg("Hauts-de-France")

# 3 Plot credits
par(mar = c(0, 0, 0, 0))
plot(0, 0, xlab = "", ylab = "", type = "n", axes = FALSE, xlim = c(-0.1, 1), ylim = c(-1, 1))
text(0, 0, "Data:
https://www.data.gouv.fr/fr/datasets/
donnees-de-laboratoires-pour-le-
depistage-indicateurs-sur-les-mutations/

Code: https://github.com/flodebarre/
nouveauCriblage/blob/main/scripts/criblage2.Rmd
", adj = 0, cex = 0.6, family = "mono")

# 4 NOR
par(mar = themar)
plotReg("Normandie")

# 5 IDF
# 
#layout(1)
plotReg("Île-de-France")

# 6 GE
plotReg("Grand Est")

# 7 BRE
plotReg("Bretagne")

# 8 CVL
plotReg("Centre-Val de Loire")

# 9 BFC
plotReg("Bourgogne-Franche-Comté")

# 10 PDL
plotReg("Pays de la Loire")

# 11 ARA
plotReg("Auvergne-Rhône-Alpes")

# 12 PACA
plotReg("Provence-Alpes-Côte d'Azur")

# 13 NAQ
plotReg("Nouvelle-Aquitaine")

# 14 OCC
plotReg("Occitanie")

# 15 COR
plotReg("Corse")
```

Alt text: graphique à 13 sous-figures, représentant pour chaque région de France métropolitaine la proportion de L452R+ au cours du temps. Augmentation partout, mais pas depuis le même moment ni exactement au même taux

tweet quote logit: https://twitter.com/flodebarre/status/1525063184150822912?s=20&t=iuVJ996CxQCT6v_UlfGh4Q

## By departement

```{r, eval = FALSE}
# Define function to plot a specific region reg
plotDep <- function(dep, xmin = "2022-03-01", addNotes = FALSE, ymax = 0.5, ymin = 5*10^-4, withMut = TRUE, col = colLiss, scale = "logit", percent = FALSE, ...){
  # withMut: whether to plot C1 / (C1 + C0) (TRUE), or C0 / (C1 + C0)
  
  # Subset of the data for this region
  dat <- dat.Deps[which(dat.Deps$dep == dep & dat.Deps$date >= xmin), ]

  #suffix <- ifelse(deliss, ".dl", "")
  id <- ifelse(withMut, 1, 0)
  
  # Plot data as shared on data.gouv.fr
  suffix <- ".sum7_mid"
  plotMut0(dat$date, dat[, paste0("nb_C", toString(id), suffix)], dat[, paste0("nb_C", toString(1 - id), suffix)], col = colLiss, thetit = "", 
           ymin = ymin, ymax = ymax, 
           addNotes = addNotes, scale = scale, percent = percent,...)
  
  # Plot data delisse
  suffix <- ".dl"
  plotMut0(dat$date, dat[, paste0("nb_C", toString(id), suffix)], dat[, paste0("nb_C", toString(1 - id), suffix)], col = colDeliss, thetit = "", 
           ymin = ymin, ymax = ymax, 
           addNotes = addNotes, add = TRUE, pch = 1, scale = scale, ...)

  # Region name as title  
  title(main = paste0(unique(dat$nom_dep), " (", dep, ")"))
}

# Check function with IDF region
layout(1)
plotDep("35")
plotDep("03")
plotDep("92")
plotDep("69")
plotDep("75")
plotDep("11")
plotDep("57")



```

Mega map

```{r, fig.width=9, fig.height=15}
setwd("../../data")
source("layout_departements_new.R")
setwd("../nouveauCriblage/scripts/")

# Plot layout, placing each departement approximately where it is on the French map
layout(deps.indices, heights = heights.deps, widths = widths.deps)

i <- 1
xmin <- "2022-03-01"
for(i in geo.departements$index){
  dep <- geo.departements[i, "code_departement"]
  
    # Subset of the data for this region
  dat <- dat.Deps[which(dat.Deps$dep == dep & dat.Deps$date >= xmin), ]

  # Plot data as shared on data.gouv.fr
  id <- 1
  suffix <- ".sum7_mid"
  par(mar = c(0, 0, 1, 0))
  plotMut0(dat$date, dat[, paste0("nb_C", toString(id), suffix)], dat[, paste0("nb_C", toString(1 - id), suffix)], col = colLiss, thetit = "", 
           ymin = 0.0001, ymax = 0.5, 
           addNotes = FALSE, scale = "logit", percent = FALSE, title = FALSE, axes = FALSE, add = FALSE)
  mtext(side = 3, text = dep, line = 0)
}
```

